---
title: "Creating mixed groups"
subtitle: <center><font size="4"><b>Group flight coordination in Thyroptera</b></font></center>
author: <center><font size="4"><a href="https://marce10.github.io/">Marcelo Araya-Salas, PhD</a> & <a href="https://www.batcr.com/">Gloriana Chaverri, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

### Define file with group and individual ID data
```{r, eval = TRUE}

# add file name including path
file.name <- "~/Downloads/Combinaciones grupos.xlsx"

# file.exists(file.name)

fi <- file.info(file.name)

replications <- 100
```

- The name of the data file is **"`r basename(file.name)`"** which is found in **"`r dirname(file.name)`"**

- The file was modified on **`r substr(x = fi$mtime, 0, 10)`** at **`r substr(x = fi$mtime, 12, 16)`**

```{r packages, message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE, include = FALSE}

## add 'developer/' to packages to be installed from github
x <- c("pbapply", "parallel","readxl", "knitr", "kableExtra")

aa <- lapply(x, function(y) {
  
  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]
  
  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  devtools::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

```

```{r test with simulated data, eval = FALSE, include=FALSE}


g1 <- c("M1-g1", "M2-g1", "F1-g1", "F2-g1", "M3-g1")
g2 <- c("M1-g2", "M2-g2", "F1-g2", "F2-g2", "F3-g2")
g3 <- c("M1-g3", "M2-g3", "F1-g3", "F2-g3")
g4 <- c("M1-g4", "F1-g4", "F2-g4", "F3-g4")
g5 <- c("M1-g5", "M2-g5", "F1-g5")
g6 <- c("M1-g6", "F1-g6")

df <- data.frame(indiv = c(g1, g2, g3, g4, g5, g6))
df$sex <- substr(0, 1, x = df$indiv)
df$group <- substr(4, 5, x = df$indiv)

cols <- c("nar", "ama", "ver", "ros")

combis <- pblapply(1:10000, cl = 3, function(y){
  
out2 <- lapply(unique(df$group), function(x){
  
  X <- df[df$group == x, ]
  X$color <- c(sample(cols, nrow(X) - 1), "sin")
  
  return(X)
})
  
df <- do.call(rbind, out2)  
  
df$indiv <- paste(df$indiv, df$color, sep = "-")

mixed_groups <- list()

df2 <- df

while(nrow(df2) > 2){

  df2 <- df2[sample(1:nrow(df2)), ]
    
  mixed_groups[[length(mixed_groups) +1]]  <-  df2$indiv[!duplicated(df2$group)]

if (length(mixed_groups[[length(mixed_groups)]]) > 5){
  cls_mg <- substr(7, 9, x = mixed_groups[[length(mixed_groups)]])
  tb <- table(cls_mg)
    if (any(names(tb)[tb > 1] == "sin") & sum(tb > 1) > 1)
        tb <- tb[names(tb) != "sin"]
  
  if (any(tb > 1))
    mixed_groups[[length(mixed_groups)]] <- mixed_groups[[length(mixed_groups)]][-1 * which(cls_mg %in% names(tb)[tb > 1])[1:(length(mixed_groups[[length(mixed_groups)]]) - 5)]] else
      mixed_groups[[length(mixed_groups)]] <- sample(mixed_groups[[length(mixed_groups)]], 5)
    }

df2 <- df2[!df2$indiv %in% mixed_groups[[length(mixed_groups)]], ]
  
}


sex_ratio <- sapply(mixed_groups, function(x) {
  
  
  sx <- table(substr(0, 1, x = x))
  
  sx["F"] / sx["M"]

})

col_count <- sapply(mixed_groups, function(x) {
  
  cl <- table(substr(7, 9, x = x))

  return(max(cl))
  
})

dup_col <- sapply(mixed_groups, function(x) {
  
  cl <- table(substr(7, 9, x = x))

  if (sum(cl > 1) == 1 & all(names(cl)[cl > 1] == "sin"))
  return("sin") else return("none")
  
})


out <- list(sex_ratio = mean(sex_ratio), left = nrow(df) - sum(sapply(mixed_groups, length)), group = mixed_groups, col_count = sum(col_count) - length(col_count), dup_col = ifelse(any(dup_col == "sin"), "sin", NA))

return(out)
})

names(combis) <- 1:length(combis)

res <- data.frame(comb = names(combis), sex_ratio = sapply(combis, "[[", "sex_ratio"), left = sapply(combis, "[[", "left"), col_count = sapply(combis, "[[", "col_count"), dup_col = sapply(combis, "[[", "dup_col"))

res <- res[res$left == 0, ]
nrow(res)
res <- res[!is.na(res$sex_ratio), ]
nrow(res)
res$abs_ratio <- abs(res$sex_ratio - 1)
res <- res[order(res$abs_ratio), ]
res <- res[!is.na(res$dup_col), ]
nrow(res)
res <- res[res$col_count <= 2, ]
nrow(res)
 
head(res)

combis[[res$comb[1]]]$group

combis[[res$comb[2]]]$group

```

</br>


---

#### Data supplied: 

Must have three columns: "Grupo", "ID" and "Sexo"

```{r read data, eval = TRUE}

df <- read_excel(file.name, col_types = c("text", "text", "text"))

df <- df[order(df$Grupo), ]

kb <-  kbl(df)  

kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

Contains data for `r length(unique(df$Grupo))` groups and `r nrow(df)` individuals

---

```{r run searching algorithm, eval = TRUE, include=FALSE}

df$indiv <- sapply(df$ID, function(x) substr(nchar(x) - 2, nchar(x), x = x))

df$sex <- toupper(df$Sexo)
df$indiv <- paste0(df$sex, df$indiv)
df$group <- substr(0, 3, x = df$Grupo)

cols <- c("nar", "ama", "ver", "ros")

pboptions(type = "none")

combis <- pblapply(1:replications, cl = 1, function(y){
  
  out2 <- lapply(unique(df$group), function(x){
    
    X <- df[df$group == x, ]
    
    if (nrow(X) <= 5) 
    X$color <- c(sample(cols, nrow(X) - 1), "sin") else
    X$color <- c(cols, "sin", paste0("mezcla", 1:(nrow(X) - 5)))
    
    return(X)
  })
    
  df <- do.call(rbind, out2)  
    
  df$indiv <- paste(df$indiv, df$color, sep = "-")
  
  mixed_groups <- list()
  
  df2 <- df
  
  while(nrow(df2) > 1){
  
    df2 <- df2[sample(1:nrow(df2)), ]
      
    mixed_groups[[length(mixed_groups) +1]]  <-  df2$indiv[!duplicated(df2$group)]
  
  if (length(mixed_groups[[length(mixed_groups)]]) > 5){
    cls_mg <- substr(7, 9, x = mixed_groups[[length(mixed_groups)]])
    tb <- table(cls_mg)
      if (any(names(tb)[tb > 1] == "sin") & sum(tb > 1) > 1)
          tb <- tb[names(tb) != "sin"]
    
    if (any(tb > 1))
      mixed_groups[[length(mixed_groups)]] <- mixed_groups[[length(mixed_groups)]][-1 * which(cls_mg %in% names(tb)[tb > 1])[1:(length(mixed_groups[[length(mixed_groups)]]) - 5)]] else
        mixed_groups[[length(mixed_groups)]] <- sample(mixed_groups[[length(mixed_groups)]], 5)
      }
  
  df2 <- df2[!df2$indiv %in% mixed_groups[[length(mixed_groups)]], ]
    
  }
  
  
  sex_ratio <- sapply(mixed_groups, function(x) {
    
    sx <- table(substr(0, 1, x = x))
    
    sx["H"] / sx["M"]
  
  })
  
  col_count <- sapply(mixed_groups, function(x) {
    
    cl <- table(substr(6, 8, x = x))
  
    return(max(cl))
    
  })
  
  dup_col <- sapply(mixed_groups, function(x) {
    
    cl <- table(substr(6, 8, x = x))
  
    if (sum(cl > 1) == 1 & all(names(cl)[cl > 1] == "sin"))
    return("sin") else return("none")
    
  })
  
  
  out <- list(sex_ratio = mean(sex_ratio), left = nrow(df) - sum(sapply(mixed_groups, length)), group = mixed_groups, col_count = sum(col_count) - length(col_count), dup_col = ifelse(any(dup_col == "sin"), "sin", NA), df = df)
  
  return(out)
})

names(combis) <- 1:length(combis)

res <- data.frame(comb = names(combis), sex_ratio = sapply(combis, "[[", "sex_ratio"), left = sapply(combis, "[[", "left"), col_count = sapply(combis, "[[", "col_count"), dup_col = sapply(combis, "[[", "dup_col"))

res <- res[res$left == min(res$left), ]

# remove those with groups of only one sex
res1 <- res[!is.na(res$sex_ratio), ]
if (nrow(res1) > 0)  res <- res1

# calculate and order by diference to 1 sex ratio
res$abs_ratio <- abs(res$sex_ratio - 1)
res <- res[order(res$abs_ratio), ]

# remove those with duplicated colors
res1 <- res[!is.na(res$dup_col), ]
if (nrow(res1) > 0)  res <- res1

# keep those with less duplicated "without color"
res1 <- res[res$col_count == min(res$col_count), ]
if (nrow(res1) > 0)  res <- res1

# selected mixed groups
sel_grp <- combis[[res$comb[1]]]$group

# original data
sel_db <- combis[[res$comb[1]]]$df

# extract selected groups
sel_df <- data.frame(grp_col = unlist(sel_grp))

sel_df$col <- substr(x = sel_df$grp_col, 6, 8)
sel_df$indiv <- substr(x = sel_df$grp_col, 0, 4)
sel_df$Grupo_mixto <- unlist(lapply(1:length(sel_grp), function(x) rep(x, length(sel_grp[[x]]))))

# merge to original data
df_col <-merge(df, sel_db[, c("ID")], all.x = TRUE)

df_col <- merge(df_col, sel_df[, c("indiv", "col", "Grupo_mixto")], all.x = TRUE)


# add label to new mixed groups
# df_col$Grupo_mixto <- sapply(df_col$indiv, function(x) {
#   y <- sel_df$Grupo_mixto[sel_df$indiv == x]
#   if (length(y) == 0) y <- "No asignado"
#   return(y)
#   })

df_col$Grupo_mixto[is.na(df_col$Grupo_mixto)] <- "No asignado"

# order by original group
df_col <- df_col[order(df_col$Grupo), ]


names(df_col)[names(df_col) == "col"] <- "Color"

df_col$Color <- gsub("ama", "Amarillo", df_col$Color)
df_col$Color <- gsub("ros", "Rosado", df_col$Color)
df_col$Color <- gsub("ver", "Verde", df_col$Color)
df_col$Color <- gsub("nar", "Naranja", df_col$Color)
df_col$Color <- gsub("sin", "Sin pintar", df_col$Color)

df_col$cols <- df_col$Color
df_col$cols <- gsub("Sin pintar", "white", df_col$cols)
df_col$cols <- gsub("Naranja", "orange", df_col$cols)
df_col$cols <- gsub("Verde", "green", df_col$cols)
df_col$cols <- gsub("Rosado", "pink", df_col$cols)
df_col$cols <- gsub("Amarillo", "yellow", df_col$cols)
df_col$cols[is.na(df_col$cols)] <- "white"
df_col$Color[is.na(df_col$Color)] <- ""

```

### Colors for individuals order by original groups

```{r colors for individuals order by original groups, eval = TRUE}

df_col <- df_col[order(df_col$Grupo), ]

kb <-  kbl(df_col[, c("Grupo", "ID", "Sexo", "Color")], row.names = FALSE)  

kb <- column_spec(kb, 4, color = "black",
              background = df_col$cols)

kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


### Colors for individuals order by mixed groups

```{r colors for individuals order by mixed groups, eval = TRUE}

df_col <- df_col[order(df_col$Grupo_mixto), ]

df_col_l <- lapply(unique(df_col$Grupo_mixto), function(x){
  
  X <- df_col[df_col$Grupo_mixto == x, ]
  X$comments <- ""
  X$comments[X$Color == "Sin pintar" & duplicated(X$Color)] <- "Debe ser coloreado"
    
  return(X)
})

df_col <- do.call(rbind, df_col_l)

kb <-  kbl(df_col[, c("Grupo_mixto", "ID", "Sexo", "Color", "comments")], row.names = FALSE)  

kb <- column_spec(kb, 4, color = "black",
              background = df_col$cols)

kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


