---
title: Inquiry calling rate in group and solo flights
subtitle: Group flight coordination in Thyroptera
author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a> & <a href="https://www.batcr.com/">Gloriana Chaverri, PhD</a>
date: "`r Sys.Date()`"
toc: true
toc-depth: 2
toc-location: left
number-sections: true
highlight-style: pygments
format:
  html:
    df-print: kable
    code-fold: true
    code-tools: true
    css: extra.css
---

```{=html}
<style>
body
  { counter-reset: source-line 0; }
pre.numberSource code
  { counter-reset: none; }
</style>
```

```{r set root directory, echo = FALSE}

# set working directory as project directory or one directory above,
rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
if (is(rootdir, "try-error")) rootdir <-  ".."
knitr::opts_knit$set(root.dir = rootdir)

```

```{r add link to github repo, echo = FALSE, results='asis'}

# print link to github repo if any
if (file.exists("./.git/config")){
  config <- readLines("./.git/config")
  url <- grep("url",  config, value = TRUE)
  url <- gsub("\\turl = |.git$", "", url)
  cat("\nSource code and data found at [", url, "](", url, ")", sep = "")
  }

```

## Load packages
```{r packages, message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE}

# add 'developer' to packages to be installed from github
x <- c("data.table", "lubridate", "devtools", github = "maRce10/warbleR", "readxl", "ranger", "caret", "e1071", "pbapply", "viridis", "ggplot2", "rlang", "Sim.DiffProc", "soundgen", "brms", "ggrepel", "posterior", "ggridges", "tidybayes", "purrr", bioconductor = "multtest", "umap", github = "maRce10/PhenotypeSpace", "metap"#, "markovchain", "igraph", "TraMineR", "spgs",
       )

source("~/Dropbox/R_package_testing/sketchy/R/load_packages.R")
load_packages(x, quite = TRUE)

```

## Functions and global parameters
```{r functions and parameters, eval = TRUE, echo = TRUE}

warbleR_options(wl = 300, parallel = 1, bp = "frange", fast = TRUE, threshold = 15, ovlp = 20)

opts_knit$set(root.dir = "..")


theme_set(theme_classic(base_size = 24))

# set chunk options
opts_chunk$set( fig.width = 7, fig.height = 4, warning = FALSE, message = FALSE, tidy = FALSE)

cols <- viridis(10, alpha = 0.7)

source("~/Dropbox/R_package_testing/brmsish/R/extended_summary.R")
source("~/Dropbox/R_package_testing/brmsish/R/contrasts.R")
source("~/Dropbox/R_package_testing/brmsish/R/helpers.R")

# model parameters
chains <- 4
iter <- 10000
priors <- c(prior(normal(0, 10), "b"), prior(normal(0, 50), "Intercept"),
    prior(student_t(3, 0, 20), "sd"), prior(student_t(3, 0, 20), "sigma"))

```

## Read data
```{r read data, eval = TRUE}

clls <- readRDS("./data/processed/curated_extended_selection_table_inquiry_calls_2020_&_2021.RDS")

sub_clls <- clls[!duplicated(clls$org.sound.files), ]
sub_clls$file.duration <- sub_clls$sound.file.samples / (sub_clls$sample.rate * 1000)
# 
# metadat2 <- read.csv("./data/processed/metadata_inquiry_calls_2020_&_2021.csv", stringsAsFactors = FALSE)

metadat <- as.data.frame(read_excel("./data/raw/Anexo 1_Proyecto MPI enero 20-21.xlsx"))
names(metadat) <- gsub(" ", ".", names(metadat))
metadat$year <- substr(metadat$Día, 0, 4)

metadat$year[is.na(metadat$year)] <- "2021"
metadat <- metadat[metadat$tipo.de.video != "calibracion de video", ]

# audio solamente para identificacion de sonidos. TASA NO
metadat <- metadat[metadat$Audio != "60 y 61", ]

metadat$year.audio <- paste(metadat$year, metadat$Audio, sep = "-")

caps <- as.data.frame(read_excel("./data/raw/Proyecto MPI enero 2020_2.xlsx", sheet = "Capturas"))

# read acoustic parameter data
acous_param_l <- readRDS("./data/processed/acoustic_parameters_all_groups_specific_warbler_acoustic_measurements_curated_data_2020_&_2021.RDS")

# read as RDS
agg_pred <- readRDS("./data/processed/predicted_individual_in_group_flights_2020_&_2021.RDS")

# read diagnostics
diagnostics <- readRDS("./data/processed/random_forests_diagnostics_solo_flight.RDS")

```

## Explore data
```{r results on predicted group flights, eval = TRUE}

# remove groups with individuals with no solo flight

diagnostics$all_with_solo <- sapply(1:nrow(diagnostics), function(x) 
{
    Y <- acous_param_l[[diagnostics$group[x]]]
  Y$sound.files <- sapply(Y$sound.files, function(x) strsplit(x, ".wav")[[1]][1])
  Y$year.audio <- sapply(Y$sound.files, function(x) sub_clls$year.audio[gsub(".wav", "", sub_clls$org.sound.files) == x][1])
    
  indivs <- strsplit(na.omit(Y$indiv[Y$year.audio == diagnostics$group_flight_files[x]])[1], split = "\\|")[[1]]

  call_per_indiv <- sapply(indivs, function(y) sum(clls$Individuo == y & clls$Experimento == "vuelo solo"))
  
  if (any(call_per_indiv == 0)) return("missing solo")  else 
    return("OK")
  }
)

diagnostics <- diagnostics[diagnostics$all_with_solo == "OK", ]

agg_pred <-agg_pred[agg_pred$group %in% diagnostics$group, ]

# get summary by group
summary_by_group_list <- lapply(unique(agg_pred$group), function(x) {
    # print(x)
  Y <- agg_pred[agg_pred$group == x, ]

#total number of calls above lowest group threshold for true positives
  above_threshold_calls <- sum(Y$max_prob > diagnostics$min_prob_threshold[diagnostics$group == x])

#proportion of calls
  prop_above_calls <- above_threshold_calls / nrow(Y)

  return(data.frame(group = x, experiment = diagnostics$experiment[diagnostics$group == x], total_calls = nrow(Y), above_threshold_calls = above_threshold_calls, prop_above_calls, n_individuals = diagnostics$n_indiv[diagnostics$group == x]))
  
})

# calls per individual and group
calls_by_indiv_group <- aggregate(sound.files ~ group + pred_indiv, agg_pred, length)

names(calls_by_indiv_group) <- c("Group", "Individual_ID", "Number_of_calls")

calls_by_indiv_group <- calls_by_indiv_group[order(calls_by_indiv_group$Group), ]

write.csv(calls_by_indiv_group, "./data/processed/calls_by_indiv_group.csv", row.names = FALSE)

summary_by_group <- do.call(rbind, summary_by_group_list)

groups_by_cat <- aggregate(group ~ experiment, summary_by_group, FUN = function(x) length(unique(x)))

groups_by_cat

```

Group count by number of individuals and experiment
```{r results on predicted group flights by experiment 2, eval = TRUE}

groups_by_cat_n <- as.data.frame.matrix(table(summary_by_group$experiment, summary_by_group$n_individuals))

groups_by_cat_n

```

---

# Stats
## Call rate

### Format data
```{r individual vs group call, eval = FALSE, echo = TRUE, fig.width = 12, fig.height = 8}

# get group and solo call counts per individual, call rate in call/min
call_rate_list <- pblapply(1:nrow(diagnostics), function(x){

  Y <- acous_param_l[[diagnostics$group[x]]]
  Y$sound.files <- sapply(Y$sound.files, function(x) strsplit(x, ".wav")[[1]][1])
  Y$year.audio <- sapply(Y$sound.files, function(x) sub_clls$year.audio[gsub(".wav", "", sub_clls$org.sound.files) == x][1])
  
  indivs_in_group <- strsplit(Y$indiv[Y$year.audio == diagnostics$group_flight_files[x]], split = "\\|")[[1]] 
  
  calls_per_sound_file_list <- lapply(na.omit(c(indivs_in_group, "group")), function(y){ 
    
      if (y != "group"){
      tab <- table(Y$year.audio[Y$indiv == y])
      if (length(tab) == 0) {tab <- 0
      names(tab) <- y}
      } else{
        tab <- sum(Y$year.audio == diagnostics$group_flight_files[x], na.rm = TRUE)
    names(tab) <- diagnostics$group_flight_files[x]
        }
    
    df <- data.frame(group = diagnostics$group[x], experiment = diagnostics$experiment[x], year.audio = names(tab), indiv = y, calls = as.vector(tab), flight.time = sapply(names(tab), USE.NAMES = FALSE, function(x) (round(as.numeric(metadat$Tiempo.de.vuelo[metadat$year.audio == x][1]) * 60 / 0.04166667))), file.duration = sapply(names(tab), function(x) (sub_clls$file.duration[sub_clls$year.audio == x][1])))
    
    return(df)
    }
  )
  
  calls_per_sound_file <- do.call(rbind, calls_per_sound_file_list)
  
  calls_per_sound_file$rate <- calls_per_sound_file$calls / calls_per_sound_file$file.duration * 60 # calls 

  return(calls_per_sound_file)
    })

call_rate_by_group <- do.call(rbind, call_rate_list)

rownames(call_rate_by_group) <- 1:nrow(call_rate_by_group)

#  add call rate to those missing
call_rate_by_group$rate <- ifelse(!is.na(call_rate_by_group$rate), call_rate_by_group$rate, call_rate_by_group$calls / call_rate_by_group$flight.time)

# add sex
call_rate_by_group$sex <- sapply(call_rate_by_group$indiv, function(x) if(x != "group") na.exclude(caps$Sexo[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

call_rate_by_group$sex <- ifelse(call_rate_by_group$sex == "m", "Male", "Female")

# add age
call_rate_by_group$age <- sapply(call_rate_by_group$indiv, function(x) if(x != "group") na.exclude(caps$Edad[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

call_rate_by_group$age <- ifelse(call_rate_by_group$age == "sa", "Sub-adult", "Adult")

# reproductive stage
call_rate_by_group$reprod.stg <- sapply(call_rate_by_group$indiv, function(x) if(x != "group") na.exclude(caps$`Estado reproductivo`[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

call_rate_by_group$reprod.stg[call_rate_by_group$reprod.stg == "p?"] <- "Pregnant"
call_rate_by_group$reprod.stg[call_rate_by_group$reprod.stg == "ne"] <- "Inactive"

call_rate_by_group$group.size <-sapply(call_rate_by_group$group, function(x) diagnostics$n_indiv[diagnostics$group == x])

# order levels
call_rate_by_group$experiment <- factor(call_rate_by_group$experiment, levels = c("regular", "mixed"))

call_rate_by_group$type <- factor(ifelse(call_rate_by_group$indiv == "group", "group", "solo"), levels = c("solo", "group"))

saveRDS(call_rate_by_group, "./data/processed/call_rate_by_group.RDS")

```

```{r calling rate per individual solo vs group, eval = FALSE, echo = TRUE, fig.width = 12, fig.height = 8}

# get group and solo call counts per individual, call rate in call/min
call_rate_indiv_list <- pblapply(1:nrow(diagnostics), function(x){

  # print(x)
  Y <- acous_param_l[[diagnostics$group[x]]]
  
  Y$pred.id <- NA
  # add id for group flights
  Y$pred.id[Y$experiment != "vuelo solo"] <- sapply(Y$sound.files[Y$experiment != "vuelo solo"], function(x) agg_pred$pred_indiv[agg_pred$sound.files == x])
  
  Y$sound.files <- sapply(Y$sound.files, function(x) strsplit(x, ".wav")[[1]][1])
  Y$year.audio <- sapply(Y$sound.files, function(x) sub_clls$year.audio[gsub(".wav", "", sub_clls$org.sound.files) == x][1])
  
  indivs_in_group <- na.omit(strsplit(Y$indiv[Y$year.audio == diagnostics$group_flight_files[x]], split = "\\|")[[1]]) 
  
  rate_indiv_list <- lapply(indivs_in_group, function(y){ 
    # print(y)
    year.audio.group <- na.omit(unique(Y$year.audio[Y$experiment != "vuelo solo"]))
    year.audio.solo <- na.omit(unique(Y$year.audio[Y$indiv == y & Y$experiment == "vuelo solo"]))
    
    flight.time.group <-  round(as.numeric(metadat$Tiempo.de.vuelo[metadat$year.audio == year.audio.group][1]) * 60 / 0.04166667)
  
    flight.time.solo <- sapply(year.audio.solo, function(w) round(as.numeric(metadat$Tiempo.de.vuelo[metadat$year.audio == w][1]) * 60 / 0.04166667)
)  
    file.duration.group <- sub_clls$file.duration[sub_clls$year.audio == year.audio.group][1]
    file.duration.solo <- sapply(year.audio.solo, function(w) sub_clls$file.duration[sub_clls$year.audio == w][1])
    
    calls.solo <- sapply(year.audio.solo, function(w) sum(Y$year.audio == w))
    calls.group <- sum(Y$year.audio == year.audio.group & Y$pred.id == y)
    
    df <- data.frame(indiv = y, group = diagnostics$group[x], size = length(indivs_in_group), experiment = diagnostics$experiment[x], type = c("group", "solo"), year.audio = c(year.audio.group, paste(year.audio.solo, collapse = "/")), calls = c(calls.group, sum(calls.solo)), flight.time = c(flight.time.group, sum(flight.time.solo)), file.duration = c(file.duration.group, sum(file.duration.solo)))
    
    return(df)
    }
  )
  
  rate_indiv <- do.call(rbind, rate_indiv_list)
  
  rate_indiv$rate <- rate_indiv$calls / rate_indiv$file.duration * 60 # calls 

  return(rate_indiv)
    })

call_rate_indiv <- do.call(rbind, call_rate_indiv_list)

rownames(call_rate_indiv) <- 1:nrow(call_rate_indiv)

#  add call rate to those missing
call_rate_indiv$rate <- ifelse(!is.na(call_rate_indiv$rate), call_rate_indiv$rate, call_rate_indiv$calls / call_rate_indiv$flight.time)

# add sex
call_rate_indiv$sex <- sapply(call_rate_indiv$indiv, function(x) if(x != "group") na.exclude(caps$Sexo[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

call_rate_indiv$sex <- ifelse(call_rate_indiv$sex == "m", "Male", "Female")

# add age
call_rate_indiv$age <- sapply(call_rate_indiv$indiv, function(x) if(x != "group") na.exclude(caps$Edad[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

call_rate_indiv$age <- ifelse(call_rate_indiv$age == "sa", "Sub-adult", "Adult")

# reproductive stage
call_rate_indiv$reprod.stg <- sapply(call_rate_indiv$indiv, function(x) if(x != "group") na.exclude(caps$`Estado reproductivo`[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

call_rate_indiv$reprod.stg[call_rate_indiv$reprod.stg == "p?"] <- "Pregnant"
call_rate_indiv$reprod.stg[call_rate_indiv$reprod.stg == "ne"] <- "Inactive"

call_rate_indiv$group.size <-sapply(call_rate_indiv$group, function(x) diagnostics$n_indiv[diagnostics$group == x])

# order levels
call_rate_indiv$experiment <- factor(call_rate_indiv$experiment, levels = c("regular", "mixed"))

call_rate_indiv$type <- factor(ifelse(call_rate_indiv$type == "group", "group", "solo"), levels = c("solo", "group"))

saveRDS(call_rate_indiv, "./data/processed/call_rate_by_individual.RDS")

```

```{r individual vs group call plot, eval = TRUE, echo = TRUE, fig.width = 12, fig.height = 8}

call_rate_by_group <- readRDS("./data/processed/call_rate_by_group.RDS")
call_rate_indiv <- readRDS("./data/processed/call_rate_by_individual.RDS")
call_rate_by_group <- call_rate_by_group[call_rate_by_group$type != "solo", ]
call_rate_by_group$type <- "overall.group"

call_rate_indiv$type <- as.character(call_rate_indiv$type)
call_rate_indiv$type[call_rate_indiv$type == "group"] <- "indiv.group" 

call_rate <- rbind(call_rate_by_group[, intersect(names(call_rate_by_group), names(call_rate_indiv))
], call_rate_indiv[, intersect(names(call_rate_by_group), names(call_rate_indiv))
])

# aggregate for plot
agg_rate <- aggregate(rate ~ experiment + type + group.size, data = call_rate, FUN = mean)

agg_rate$sd <- aggregate(rate ~ experiment + type + group.size, data = call_rate, FUN = sd)$rate

agg_rate$type <- factor(agg_rate$type, levels = c("solo", "indiv.group", "overall.group"))

# ggplot(agg_rate[agg_rate$group.size < 6, ], aes(x = type, y = rate)) + 
#   geom_point(color = viridis(10)[3], size = 2) +
#     geom_errorbar(aes(ymin = rate - sd, ymax = rate + sd), color = viridis(10)[3], width=.1, lwd = 1.1) +
#    geom_hline(yintercept = mean(call_rate$rate[call_rate$type == "solo"], na.rm = TRUE), lty = 2, alpha = 0.5) + 
#   labs(x = "Flight", y = "Call rate (calls / min)") + 
#   facet_grid(group.size ~ experiment) + 
#     theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Individual call rate in solo and group flight
```{r call rate per individual solo vs group lines plot, eval = TRUE, echo = TRUE, fig.width = 12, fig.height = 8}

# aggregate for plot
agg_rate_indiv <- aggregate(rate ~ experiment + type + group.size + indiv, data = call_rate_indiv, FUN = mean)

agg_rate_indiv$sd <- aggregate(rate ~ experiment + type + group.size + indiv, data = call_rate_indiv, FUN = sd)$rate

agg_rate_indiv <- agg_rate_indiv[agg_rate_indiv$indiv != "group", ]



mixed_ids <- unique(agg_rate_indiv$indiv[agg_rate_indiv$experiment == "mixed"])

mixed_l <- lapply(mixed_ids, function(x) {
  
  X <- agg_rate_indiv[agg_rate_indiv$indiv == x, ]
  X <- X[X$experiment == "mixed" | X$type == "solo", ]
  
  if (sum(X$experiment == "mixed") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "mixed"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

mixed <- do.call(rbind, mixed_l)

mixed$experiment <- "mixed"

# regular
regular_ids <- unique(agg_rate_indiv$indiv[agg_rate_indiv$experiment == "regular"])

regular_l <- lapply(regular_ids, function(x) {
  
  X <- agg_rate_indiv[agg_rate_indiv$indiv == x, ]
  X <- X[X$experiment == "regular" | X$type == "solo", ]
  
  if (sum(X$experiment == "regular") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "regular"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

regular <- do.call(rbind, regular_l)

regular$experiment <- "regular"

sub_agg <- rbind(mixed, regular)

sub_agg$group <- sapply(seq_len(nrow(sub_agg)), function(x){
  out <- if (sub_agg$type[x] == "Solo")
    NA else
      unique(metadat$Grupo[metadat$Individuo == sub_agg$indiv[x] & grepl(if (sub_agg$experiment[x] == "regular") "sin sonido" else "mixto", metadat$Experimento) & !is.na(metadat$Individuo) & !is.na(metadat$Experimento)])
  return(out)
  })

sub_agg$experiment_f <- factor(ifelse(sub_agg$experiment == "regular", "Regular", "Artificial"), levels = c("Regular", "Artificial"))

sub_agg$type_f <- factor(ifelse(sub_agg$type == "solo", "Solo", "In group"), levels = c("Solo", "In group"))

# boxplots
fill_color <- adjustcolor("#e85307", 0.6)

# aggregate data for sample sizes on plot
agg_dat <- aggregate(rate ~ type_f + experiment_f, sub_agg, mean)
agg_dat$n <- sapply(1:nrow(agg_dat), function(x) length(unique(sub_agg$indiv[sub_agg$type_f == agg_dat$type_f[x] & sub_agg$experiment_f == agg_dat$experiment_f[x]]))) 
agg_dat$n.labels <- paste("n =", agg_dat$n)
agg_dat$experiment_f <- factor(agg_dat$experiment_f)
agg_dat$type_f <- factor(agg_dat$type_f)
agg_dat$n.group <-  sapply(1:nrow(agg_dat), function(x) length(unique(unlist(sub_agg$group[sub_agg$type_f == agg_dat$type_f[x] & sub_agg$experiment_f == agg_dat$experiment_f[x]]))))
agg_dat$n.labels <- ifelse(agg_dat$type_f == "In group", paste0(agg_dat$n.labels, " (groups = ", agg_dat$n.group, ")"), agg_dat$n.labels)
  
ggplot(sub_agg, aes(y = rate, x = type_f)) +
  ## add half-violin from {ggdist} package
  ggdist::stat_halfeye(
    fill = fill_color,
    alpha = 0.5,
    ## custom bandwidth
    adjust = .5,
    ## adjust height
    width = .6,
    .width = 0,
    ## move geom to the cright
    justification = -.2,
    point_colour = NA
  ) +
  geom_boxplot(fill = fill_color,
    width = .15,
    ## remove outliers
    outlier.shape = NA ## `outlier.shape = NA` works as well
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    color = fill_color,
    ## draw jitter on the left
    side = "l",
    ## control range of jitter
    range_scale = .4,
    ## add some transparency
    alpha = .5,
    transformation = ggplot2::position_jitter(height = 0)
  ) +
    labs(x="Flight", y="Call rate (calls/min)"
  ) +
   ylim(c(-1.5, 30)) +
  geom_text(data = agg_dat, aes(y = rep(-1.5, nrow(agg_dat)), x = type_f, label = n.labels), nudge_x = 0, size = 6) +
   theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
   facet_grid( ~ experiment_f)

```

```{r individual call rate stats, eval = FALSE}

call_rate_indiv <- call_rate_indiv[!is.na(call_rate_indiv$rate) & call_rate_indiv$rate > 0, ]

call_rate_indiv$experiment.type <- ifelse(call_rate_indiv$type == "solo", "solo", paste(call_rate_indiv$experiment, call_rate_indiv$type, sep = "-"))

call_rate_indiv$experiment.type <- gsub("mixed-indiv.group", "artificial.group", call_rate_indiv$experiment.type)
call_rate_indiv$experiment.type <- gsub("regular-indiv.group", "real.group", call_rate_indiv$experiment.type)

call_rate_indiv$experiment.type <- factor(call_rate_indiv$experiment.type, levels = c("solo", "real.group", "artificial.group")) 

# mean centering group size
call_rate_indiv$group.size <- call_rate_indiv$group.size - mean(call_rate_indiv$group.size)

 mod <- brm(
          formula = rate ~ experiment.type + experiment.type:group.size + (1 | indiv),
          iter = iter,
          thin = 1,
          data = call_rate_indiv,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains, 
          prior = priors, 
          file = "./data/processed/individual_call_rate_solo_vs_group"
          )

```

```{r individual call rate results, eval = TRUE, results = "asis"}

extended_summary(read.file = "./data/processed/individual_call_rate_solo_vs_group.rds", gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_", remove.intercepts = TRUE, highlight = TRUE, print.name = FALSE)

```

&nbsp;

<div class="alert alert-info">

#### Takeaways {.unnumbered .unlisted}

- Individuals call at lower rates during group flight (both real and artificial) compared to solo flight
- Rates also decreases in artificial groups as a function of group size

</div>

&nbsp;

## Individual call rate vs overall group call rate

```{r call rate per group solo vs group lines plot, eval = TRUE, echo = TRUE, fig.width = 12, fig.height = 8}

call_rate_by_group <- readRDS("./data/processed/call_rate_by_group.RDS")

call_rate_by_group$experiment.type <- ifelse(call_rate_by_group$type == "solo", "solo", paste(call_rate_by_group$experiment, call_rate_by_group$type, sep = "-"))

call_rate_by_group$experiment.type <- gsub("mixed.group", "artificial.group", call_rate_by_group$experiment.type)
call_rate_by_group$experiment.type <- gsub("regular.group", "real.group", call_rate_by_group$experiment.type)

call_rate_by_group$experiment.type <- factor(call_rate_by_group$experiment.type, levels = c("solo", "real.group", "artificial.group")) 

# aggregate for plot
agg_rate_group <- aggregate(rate ~ experiment + type + group.size + group, data = call_rate_by_group, FUN = mean)

agg_rate_group$sd <- aggregate(rate ~ experiment + type + group.size + group, data = call_rate_by_group, FUN = sd)$rate

agg_rate_group <- agg_rate_group[agg_rate_group$type != "indiv.group", ]

mixed_ids <- unique(agg_rate_group$group[agg_rate_group$experiment == "mixed"])

mixed_l <- lapply(mixed_ids, function(x) {
  
  X <- agg_rate_group[agg_rate_group$group == x, ]
  X <- X[X$experiment == "mixed" | X$type == "solo", ]
  
  if (sum(X$experiment == "mixed") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "mixed"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

mixed <- do.call(rbind, mixed_l)

mixed$experiment <- "mixed"

# regular
regular_ids <- unique(agg_rate_group$group[agg_rate_group$experiment == "regular"])

regular_l <- lapply(regular_ids, function(x) {
  
  X <- agg_rate_group[agg_rate_group$group == x, ]
  X <- X[X$experiment == "regular" | X$type == "solo", ]
  
  if (sum(X$experiment == "regular") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "regular"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

regular <- do.call(rbind, regular_l)

regular$experiment <- "regular"

sub_agg <- rbind(mixed, regular)

sub_agg$experiment_f <- factor(ifelse(sub_agg$experiment == "regular", "Regular", "Artificial"), levels = c("Regular", "Artificial"))

sub_agg$type_f <- factor(ifelse(sub_agg$type == "solo", "Solo", "Overall group"), levels = c("Solo", "Overall group"))

agg_dat$type_f <- as.character(agg_dat$type_f)
agg_dat$type_f <- gsub("In group", "Overall group", agg_dat$type_f)
agg_dat$type_f <- as.factor(agg_dat$type_f)


# composed box plot
ggplot(sub_agg, aes(y = rate, x = type_f)) +
  ## add half-violin from {ggdist} package
  ggdist::stat_halfeye(
    fill = fill_color,
    alpha = 0.5,
    ## custom bandwidth
    adjust = .5,
    ## adjust height
    width = .6,
    .width = 0,
    ## move geom to the cright
    justification = -.2,
    point_colour = NA
  ) +
  geom_boxplot(fill = fill_color,
    width = .15,
    ## remove outliers
    outlier.shape = NA ## `outlier.shape = NA` works as well
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    color = fill_color,
    ## draw jitter on the left
    side = "l",
    ## control range of jitter
    range_scale = .4,
    ## add some transparency
    alpha = .5,
    transformation = ggplot2::position_jitter(height = 0)
  ) +
    labs(x="Flight", y="Call rate (calls/min)"
  ) +
  ylim(c(-1.5, 30)) +
  geom_text(data = agg_dat, aes(y = rep(-1.5, nrow(agg_dat)), x = type_f, label = n.labels), nudge_x = 0, size = 6) +
   theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
   facet_grid( ~ experiment_f)

```

```{r individual vs group call rate stats, eval = FALSE}

call_rate_by_group <- readRDS("./data/processed/call_rate_by_group.RDS")

call_rate_by_group$experiment.type <- ifelse(call_rate_by_group$type == "solo", "solo", paste(call_rate_by_group$experiment, call_rate_by_group$type, sep = "-"))

call_rate_by_group$experiment.type <- gsub("mixed.group", "artificial.group", call_rate_by_group$experiment.type)
call_rate_by_group$experiment.type <- gsub("regular.group", "real.group", call_rate_by_group$experiment.type)

call_rate_by_group$experiment.type <- factor(call_rate_by_group$experiment.type, levels = c("solo", "real.group", "artificial.group")) 

# mean centering group size
call_rate_by_group$group.size <- call_rate_by_group$group.size - mean(call_rate_by_group$group.size)

 mod <- brm(
          formula = rate ~ experiment.type + experiment.type:group.size + (1 | group),
          iter = iter,
          thin = 1,
          data = call_rate_by_group,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains, 
          prior = priors, 
          file = "./data/processed/individual_vs_group_call_rate"
          )

```

```{r individual vs group call rate results, eval = TRUE, results = "asis"}

extended_summary(read.file = "./data/processed/individual_vs_group_call_rate.rds", gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_", remove.intercepts = TRUE, highlight = TRUE, print.name = FALSE)

```


&nbsp;

<div class="alert alert-info">

#### Takeaways {.unnumbered .unlisted}

- No difference in overall call rate between solo and group flights (both real and artificial)

</div>

&nbsp;

## Gaps 

### Individual gaps in solo vs group flights
```{r gaps per individual solo vs group, eval = FALSE, echo = TRUE, fig.width = 12, fig.height = 8}


seltab <- attributes(clls)$check.results
seltab$year.audio <- clls$year.audio
seltab <- seltab[!is.na(seltab$sound.files), ]
seltab$type <- sapply(seltab$year.audio, function(x) metadat$Experimento[metadat$year.audio == x][1])
seltab$group <- sapply(seltab$year.audio, function(x) metadat$Grupo[metadat$year.audio == x][1])
seltab <- seltab[!is.na(seltab$group), ]
seltab$type <-ifelse(grepl("solo", seltab$type), "solo", "group")
seltab$indiv <- sapply(1:nrow(seltab), function(x) if (seltab$type[x] == "solo") metadat$Individuo[metadat$year.audio == seltab$year.audio[x]] else "group")
seltab$start <- seltab$orig.start
seltab$end <- seltab$orig.end

# get group and solo call counts per individual, call rate in call/min
gaps_indiv_list <- pblapply(1:nrow(diagnostics), function(x){

  # print(x)
  # get indivs
  sound.files <- acous_param_l[[diagnostics$group[x]]]$sound.files
  
    Y <- seltab[seltab$sound.files %in% sound.files, ]

  # add id for group flights
  Y$pred.id <- NA
  Y$pred.id[Y$type == "group"] <- sapply(Y$sound.files[Y$type == "group"], function(x) agg_pred$pred_indiv[agg_pred$sound.files == x])

  Y <- Y[order(Y$orig.sound.files, Y$start), ]
    
  gaps_indiv_list <- lapply(unique(Y$indiv), function(y){ 
# print(y)
    if (y != "group"){
      W <- Y[Y$indiv == y, ]
      W$sound.files <- W$orig.sound.files
      
      gaps.solo <- data.frame(group =  diagnostics$group[x], indiv = y, experiment =  diagnostics$experiment[x], type = "solo", gaps = if (nrow(W) > 0) na.omit(gaps(W, pb = FALSE)$gaps) else NA)
    
      Z <- Y[Y$pred.id == y & !is.na(Y$pred.id), ]
      Z$sound.files <- Z$orig.sound.files 
      group.gaps <- if (nrow(Z) > 0) na.omit(gaps(Z, pb = FALSE)$gaps) else NA
      gaps.group <- data.frame(group =  diagnostics$group[x], indiv = y, experiment =  diagnostics$experiment[x], type = "indiv.group", gaps =  if (length(group.gaps) > 0) group.gaps else NA)
       
    gaps_data <- rbind(gaps.group, gaps.solo)  
    
    } else {
      Q <- Y[Y$type == "group", ]
      Q$sound.files <- Q$orig.sound.files
    gaps_data <- data.frame(group =  diagnostics$group[x], indiv = y, experiment = diagnostics$experiment[x], type = "overall.group", gaps =  na.omit(gaps(Q, pb = FALSE)$gaps))
      
return(gaps_data)
    }
  })
  gaps_indiv <- do.call(rbind, gaps_indiv_list)
  gaps_indiv$group.size <- diagnostics$n_indiv[x]
  
  
    return(gaps_indiv)
    }
  )
  
gaps_indiv <- do.call(rbind, gaps_indiv_list)
  
# add sex
gaps_indiv$sex <- sapply(gaps_indiv$indiv, function(x) if(x != "group") na.exclude(caps$Sexo[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

gaps_indiv$sex <- ifelse(gaps_indiv$sex == "m", "Male", "Female")

# add age
gaps_indiv$age <- sapply(gaps_indiv$indiv, function(x) if(x != "group") na.exclude(caps$Edad[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

gaps_indiv$age <- ifelse(gaps_indiv$age == "sa", "Sub-adult", "Adult")

# reproductive stage
gaps_indiv$reprod.stg <- sapply(gaps_indiv$indiv, function(x) if(x != "group") na.exclude(caps$`Estado reproductivo`[caps$Murci == x])[1] else NA, USE.NAMES = FALSE)

gaps_indiv$reprod.stg[gaps_indiv$reprod.stg == "p?"] <- "Pregnant"
gaps_indiv$reprod.stg[gaps_indiv$reprod.stg == "ne"] <- "Inactive"

gaps_indiv$group.size <-sapply(gaps_indiv$group, function(x) diagnostics$n_indiv[diagnostics$group == x])

# order levels
gaps_indiv$experiment <- factor(gaps_indiv$experiment, levels = c("Regular", "Artificial"))

gaps_indiv$experiment.type <- ifelse(gaps_indiv$type == "solo", "solo", paste(gaps_indiv$experiment, gaps_indiv$type, sep = "-"))


# #find individuals with repeated solo data
# indivs_2_sets <- sapply(unique(gaps_indiv$indiv), function(x) length(unique(gaps_indiv$group[gaps_indiv$indiv == x & gaps_indiv$type == "solo"])))
# 
# # remove those
# indivs_2_sets <- names(indivs_2_sets)[indivs_2_sets > 1]
# indivs_2_sets <- indivs_2_sets[indivs_2_sets != "group"]
# 
# gaps_indiv_filter_list <- lapply(unique(gaps_indiv$indiv), function(x){
# 
#   Y <- gaps_indiv[gaps_indiv$indiv == x, ]
#   
#   if(x %in% indivs_2_sets){
#     Y <- Y[Y$group == unique(Y$group)[1] & Y$type == "solo" | Y$type == "indiv.group",]
#     
#   }
#   
#   return(Y)
#   
# })

# gaps_indiv_filter <- do.call(rbind, gaps_indiv_filter_list)
# 
# # all have 1
# all(sapply(unique(gaps_indiv_filter$indiv), function(x) length(unique(gaps_indiv_filter$group[gaps_indiv_filter$indiv == x & gaps_indiv_filter$type == "solo"]))) <= 1)

 saveRDS(gaps_indiv, "./data/processed/gaps_by_individual.RDS")

```

```{r gaps per individual solo vs group plot, eval = TRUE, echo = TRUE, fig.width = 12, fig.height = 8}
gaps_indiv <- readRDS("./data/processed/gaps_by_individual.RDS")


# aggregate for plot
agg_gaps <- aggregate(gaps ~ experiment + type + group.size, data = gaps_indiv, FUN = mean)

agg_gaps$sd <- aggregate(gaps ~ experiment + type + group.size, data = gaps_indiv, FUN = sd)$gaps

```

```{r gaps per individual solo vs group lines plot, eval = TRUE, echo = TRUE, fig.width = 12, fig.height = 8}

# aggregate for plot
agg_gaps_indiv <- aggregate(gaps ~ experiment + type + group.size + indiv, data = gaps_indiv, FUN = mean)

agg_gaps_indiv$sd <- aggregate(gaps ~ experiment + type + group.size + indiv, data = gaps_indiv, FUN = sd)$gaps

agg_gaps_indiv <- agg_gaps_indiv[agg_gaps_indiv$indiv != "group", ]

mixed_ids <- unique(agg_gaps_indiv$indiv[agg_gaps_indiv$experiment == "mixed"])

mixed_l <- lapply(mixed_ids, function(x) {
  
  X <- agg_gaps_indiv[agg_gaps_indiv$indiv == x, ]
  X <- X[X$experiment == "mixed" | X$type == "solo", ]
  
  if (sum(X$experiment == "mixed") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "mixed"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

mixed <- do.call(rbind, mixed_l)

mixed$experiment <- "mixed"

# regular
regular_ids <- unique(agg_gaps_indiv$indiv[agg_gaps_indiv$experiment == "regular"])

regular_l <- lapply(regular_ids, function(x) {
  
  X <- agg_gaps_indiv[agg_gaps_indiv$indiv == x, ]
  X <- X[X$experiment == "regular" | X$type == "solo", ]
  
  if (sum(X$experiment == "regular") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "regular"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

regular <- do.call(rbind, regular_l)

regular$experiment <- "regular"

sub_agg <- rbind(mixed, regular)

sub_agg$group <- sapply(seq_len(nrow(sub_agg)), function(x){
  out <- if (sub_agg$type[x] == "Solo")
    NA else
      unique(metadat$Grupo[metadat$Individuo == sub_agg$indiv[x] & grepl(if (sub_agg$experiment[x] == "regular") "sin sonido" else "mixto", metadat$Experimento) & !is.na(metadat$Individuo) & !is.na(metadat$Experimento)])
  return(out)
  })

sub_agg$experiment_f <- factor(ifelse(sub_agg$experiment == "regular", "Regular", "Artificial"), levels = c("Regular", "Artificial"))

sub_agg$type_f <- factor(ifelse(sub_agg$type == "solo", "Solo", "In group"), levels = c("Solo", "In group"))

# aggregate data for sample sizes on plot
agg_dat <- aggregate(gaps ~ type_f + experiment_f, sub_agg, mean)
agg_dat$n <- sapply(1:nrow(agg_dat), function(x) length(unique(sub_agg$indiv[sub_agg$type_f == agg_dat$type_f[x] & sub_agg$experiment_f == agg_dat$experiment_f[x]]))) 
agg_dat$n.labels <- paste("n =", agg_dat$n)
agg_dat$experiment_f <- factor(agg_dat$experiment_f)
agg_dat$type_f <- factor(agg_dat$type_f)
agg_dat$n.group <-  sapply(1:nrow(agg_dat), function(x) length(unique(unlist(sub_agg$group[sub_agg$type_f == agg_dat$type_f[x] & sub_agg$experiment_f == agg_dat$experiment_f[x]]))))
agg_dat$n.labels <- ifelse(agg_dat$type_f == "In group", paste0(agg_dat$n.labels, " (groups = ", agg_dat$n.group, ")"), agg_dat$n.labels)
  
# composed box plot
ggplot(sub_agg, aes(y = gaps, x = type_f)) +
  ## add half-violin from {ggdist} package
  ggdist::stat_halfeye(
    fill = fill_color,
    alpha = 0.5,
    ## custom bandwidth
    adjust = .5,
    ## adjust height
    width = .6,
    .width = 0,
    ## move geom to the cright
    justification = -.2,
    point_colour = NA
  ) +
  geom_boxplot(fill = fill_color,
    width = .15,
    ## remove outliers
    outlier.shape = NA ## `outlier.shape = NA` works as well
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    color = fill_color,
    ## draw jitter on the left
    side = "l",
    ## control range of jitter
    range_scale = .4,
    ## add some transparency
    alpha = .5,
    transformation = ggplot2::position_jitter(height = 0)
  ) +
    labs(x="Flight", y = "Gap duration (s)") +
  ylim(c(-1.5, 30)) +
  geom_text(data = agg_dat, aes(y = rep(-1.5, nrow(agg_dat)), x = type_f, label = n.labels), nudge_x = 0, size = 6) +
   theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
   facet_grid( ~ experiment_f)

```

```{r gaps per individual solo vs group stats, eval = FALSE}

sub_gaps_indiv <- gaps_indiv[gaps_indiv$type != "overall.group", ]

sub_gaps_indiv$experiment.type <- gsub("-indiv.group", "", sub_gaps_indiv$experiment.type)

sub_gaps_indiv$experiment.type <- gsub("mixed", "artificial.group", sub_gaps_indiv$experiment.type)
sub_gaps_indiv$experiment.type <- gsub("regular", "real.group", sub_gaps_indiv$experiment.type)

sub_gaps_indiv$experiment.type <- factor(sub_gaps_indiv$experiment.type, levels = c("solo", "real.group", "artificial.group")) 

# mean centering group size
sub_gaps_indiv$group.size <- sub_gaps_indiv$group.size - mean(sub_gaps_indiv$group.size)

 mod <- brm(
          formula =  gaps ~ experiment.type + experiment.type:group.size + (1 | indiv),
          iter = iter,
          thin = 1,
          data = sub_gaps_indiv,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains, 
          prior = priors, 
          file = "./data/processed/individual_gaps_solo_vs_group"
          )

```

```{r gaps per individual solo vs group results, eval = TRUE, results = "asis"}

extended_summary(read.file = "./data/processed/individual_gaps_solo_vs_group.rds", gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_", remove.intercepts = TRUE, highlight = TRUE, print.name = FALSE)

# contrasts
cat("Compare artificial vs real group gaps:\n")
```

```{r gaps per individual solo vs group results contrasts, eval = TRUE, results='asis'}

# contrasts
contrasts(read.file = "./data/processed/individual_gaps_solo_vs_group.rds", predictor = "experiment.type", n.posterior = 2000, level.sep = " VS ",  html.table = TRUE, plot = TRUE, highlight = TRUE)

```

```{r gaps per individual solo vs group random slope stats DEPRECATED, eval = FALSE}

sub_gaps_indiv <- gaps_indiv[gaps_indiv$type != "overall.group", ]

sub_gaps_indiv$experiment.type <- gsub("-indiv.group", "", sub_gaps_indiv$experiment.type)

sub_gaps_indiv$experiment.type <- gsub("mixed", "artificial.group", sub_gaps_indiv$experiment.type)
sub_gaps_indiv$experiment.type <- gsub("regular", "real.group", sub_gaps_indiv$experiment.type)

sub_gaps_indiv$experiment.type <- factor(sub_gaps_indiv$experiment.type, levels = c("solo", "real.group", "artificial.group")) 

# mean centering group size
sub_gaps_indiv$group.size <- sub_gaps_indiv$group.size - mean(sub_gaps_indiv$group.size)


 mod <- brm(
          formula = gaps ~ experiment.type + experiment.type:group.size + (experiment.type | indiv),
          iter = iter,
          thin = 1,
          data = sub_gaps_indiv,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains, 
          prior = priors, 
          file = "./data/processed/individual_gaps_solo_vs_group_random_slope"
          )

```

```{r gaps per individual solo vs group random slope results DEPRECATED, eval = FALSE, results = "asis"}

extended_summary(read.file = "./data/processed/individual_gaps_solo_vs_group_random_slope.rds", gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_", remove.intercepts = TRUE, highlight = TRUE, print.name = FALSE)

# 
# # contrasts
cat("Compare artificial vs real group gaps:\n")

```

```{r gaps per individual solo vs group random slope results contrasts DEPRECATED, eval = FALSE}

# contrasts
contrasts(read.file = "../data/processed/individual_gaps_solo_vs_group_random_slope.rds", predictor = "experiment.type", n.posterior = 2000, level.sep = " VS ",  html.table = TRUE, plot = TRUE, highlight = TRUE)

```

&nbsp;

<div class="alert alert-info">

### Takeaways {.unnumbered .unlisted}

- Gaps in group flights (both mixed and regular groups) increases compare to solo flights but the magnitude of the response varies individually
- Group size also affects the magnitude of change in group flights 

</div>

&nbsp;

### Individual gaps vs overall group gaps

```{r gaps per group solo vs group lines plot, eval = TRUE, echo = TRUE, fig.width = 12, fig.height = 8}

# aggregate for plot
agg_gaps_group <- aggregate(gaps ~ experiment + type + group.size + group, data = gaps_indiv, FUN = mean)

agg_gaps_group$sd <- aggregate(gaps ~ experiment + type + group.size + group, data = gaps_indiv, FUN = sd)$gaps

agg_gaps_group <- agg_gaps_group[agg_gaps_group$type != "indiv.group", ]

mixed_ids <- unique(agg_gaps_group$group[agg_gaps_group$experiment == "mixed"])

mixed_l <- lapply(mixed_ids, function(x) {
  
  X <- agg_gaps_group[agg_gaps_group$group == x, ]
  X <- X[X$experiment == "mixed" | X$type == "solo", ]
  
  if (sum(X$experiment == "mixed") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "mixed"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

mixed <- do.call(rbind, mixed_l)

mixed$experiment <- "mixed"

# regular
regular_ids <- unique(agg_gaps_group$group[agg_gaps_group$experiment == "regular"])

regular_l <- lapply(regular_ids, function(x) {
  
  X <- agg_gaps_group[agg_gaps_group$group == x, ]
  X <- X[X$experiment == "regular" | X$type == "solo", ]
  
  if (sum(X$experiment == "regular") > 1)
    X <- X[X$group.size == max(X$group.size[X$experiment == "regular"]) | X$type == "solo",]
  
  if (nrow(X) == 1)
    X <- X[vector(), ]
    
  return(X)
  })

regular <- do.call(rbind, regular_l)

regular$experiment <- "regular"

sub_agg <- rbind(mixed, regular)

sub_agg$experiment_f <- factor(ifelse(sub_agg$experiment == "regular", "Regular", "Artificial"), levels = c("Regular", "Artificial"))

sub_agg$type_f <- factor(ifelse(sub_agg$type == "solo", "Solo", "Overall group"), levels = c("Solo", "Overall group"))

agg_dat$type_f <- as.character(agg_dat$type_f)
agg_dat$type_f <- gsub("In group", "Overall group", agg_dat$type_f)
agg_dat$type_f <- as.factor(agg_dat$type_f)

# composed box plot
ggplot(sub_agg, aes(y = gaps, x = type_f)) +
  ## add half-violin from {ggdist} package
  ggdist::stat_halfeye(
    fill = fill_color,
    alpha = 0.5,
    ## custom bandwidth
    adjust = .5,
    ## adjust height
    width = .6,
    .width = 0,
    ## move geom to the cright
    justification = -.2,
    point_colour = NA
  ) +
  geom_boxplot(fill = fill_color,
    width = .15,
    ## remove outliers
    outlier.shape = NA ## `outlier.shape = NA` works as well
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    color = fill_color,
    ## draw jitter on the left
    side = "l",
    ## control range of jitter
    range_scale = .4,
    ## add some transparency
    alpha = .5,
    transformation = ggplot2::position_jitter(height = 0)
  ) +
    labs(x="Flight", y= "Gap duration (s)") +
  ylim(c(-1.5, 10)) +
  geom_text(data = agg_dat, aes(y = rep(-1.5, nrow(agg_dat)), x = type_f, label = n.labels), nudge_x = 0, size = 6) +
   theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
   facet_grid( ~ experiment_f)

```

```{r gaps per group solo vs group stats, eval = FALSE}

sub_gaps_indiv <- gaps_indiv[gaps_indiv$type != "indiv.group", ]

sub_gaps_indiv$experiment.type <- gsub("-overall.group", "", sub_gaps_indiv$experiment.type)

sub_gaps_indiv$experiment.type <- gsub("mixed", "artificial.group", sub_gaps_indiv$experiment.type)
sub_gaps_indiv$experiment.type <- gsub("regular", "real.group", sub_gaps_indiv$experiment.type)

sub_gaps_indiv$experiment.type <- factor(sub_gaps_indiv$experiment.type, levels = c("solo", "real.group", "artificial.group")) 

 mod <- brm(
          formula =  gaps ~ experiment.type + experiment.type:group.size + (1 | group),
          iter = iter,
          thin = 1,
          data = sub_gaps_indiv,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains, 
          prior = priors, 
          file = "./data/processed/group_gaps_solo_vs_group"
 )
 
```

```{r gaps per group solo vs group results, eval = TRUE, results = "asis"}

extended_summary(read.file = "./data/processed/group_gaps_solo_vs_group.rds", gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_", remove.intercepts = TRUE, highlight = TRUE, print.name = FALSE)

# contrasts
cat("Compare artificial vs real group gaps:\n")
```

```{r gaps per group solo vs group results contrasts, eval = TRUE, results='asis'}

# contrasts
contrasts(read.file = "./data/processed/group_gaps_solo_vs_group.rds", predictor = "experiment.type", n.posterior = 2000, level.sep = " VS ",  html.table = TRUE, plot = TRUE, highlight = TRUE)

```

&nbsp;

<div class="alert alert-info">

### Takeaways {.unnumbered .unlisted}

- Gaps in artificial groups are longer than in solo flights (but not for real groups)
- Group size positively affects gap duration in both types of group flights

</div>

&nbsp;

## Time partitioning during group flight

```{r format data to measure coordination, eval = FALSE}

seltab <- attributes(clls)$check.results
seltab$year.audio <- clls$year.audio
seltab <- seltab[!is.na(seltab$sound.files), ]

seltab$type <- sapply(seltab$year.audio, function(x) metadat$Experimento[metadat$year.audio == x][1])
seltab$group <- sapply(seltab$year.audio, function(x) metadat$Grupo[metadat$year.audio == x][1])
seltab <- seltab[!is.na(seltab$group), ]
seltab$type <-ifelse(grepl("solo|Individual", seltab$type), "solo", ifelse(grepl("mixto", seltab$type), "Artificial", "Regular"))
seltab$indiv <- sapply(1:nrow(seltab), function(x) if (seltab$type[x] == "solo") metadat$Individuo[metadat$year.audio == seltab$year.audio[x]] else "group")
seltab$start <- seltab$orig.start
seltab$end <- seltab$orig.end

# get group and solo call counts per individual, call rate in call/min
pred_indiv_list <- pblapply(1:nrow(diagnostics), function(x){

  # print(x)
  # get indivs
  sound.files <- acous_param_l[[diagnostics$group[x]]]$sound.files
  
    Y <- seltab[seltab$sound.files %in% sound.files, ]
    Y <- Y[grep("solo", Y$type, invert = TRUE), ]
  # add id for group flights
  Y$pred.id <- NA
  Y$pred.id <- sapply(Y$sound.files, function(x){ 
    w <- agg_pred$pred_indiv[agg_pred$sound.files == x]
    if (length(w) == 0) w <- NA
    return(w)
    })

  Y <- Y[order(Y$orig.sound.files, Y$start), ]
  Y$start <- Y$orig.start
  Y$end <- Y$orig.end
  Y$selec <- Y$orig.selec
  Y$sing.event <- Y$orig.sound.files
  Y$indiv <- Y$pred.id
  return(Y[, c("sing.event", "indiv", "selec", "start", "end", "group", "type")])
    }
  )
  
pred_indiv_df <- do.call(rbind, pred_indiv_list)
  
sum(is.na(pred_indiv_df$indiv)) / nrow(pred_indiv_df)

unique(pred_indiv_df$x[is.na(pred_indiv_df$indiv)])


pred_indiv_df <- pred_indiv_df[!is.na(pred_indiv_df$indiv), ]

pred_indiv_df$group.size <-sapply(pred_indiv_df$group, function(x) diagnostics$n_indiv[diagnostics$group == x])

# order levels
pred_indiv_df$experiment <- pred_indiv_df$type

pred_indiv_df$experiment <- factor(pred_indiv_df$experiment, levels = c("Regular", "Artificial"))

source("~/Dropbox/R_package_testing/warbleR/R/test_coordination.R")
source("~/Dropbox/R_package_testing/warbleR/R/internal_functions.R")

# save data for damien farine
# pred_indiv_df2 <- pred_indiv_df
# pred_indiv_df2$time <- (pred_indiv_df2$start + pred_indiv_df2$end) / 2
# names(pred_indiv_df2)[1] <- "event"
# tab <- table(pred_indiv_df2$event)
# tab <- tab[tab > 19]
# write.csv(pred_indiv_df2[pred_indiv_df2$event %in% names(tab), c("group", "indiv", "event", "time", "type")], "group_call_individual_id_sequence.csv", row.names = FALSE)

coordination_results <- test_coordination(X = pred_indiv_df, iterations = 10000, less.than.chance = TRUE, ovlp.method = "time.closest", parallel = 20, cutoff = 1)

coordination_results$n.calls <- sapply(coordination_results$sing.event, function(x) sum(pred_indiv_df$sing.event == x))

coordination_results$n.calls <- sapply(coordination_results$sing.event, function(x) sum(pred_indiv_df$sing.event == x))

coordination_results$experiment <- sapply(coordination_results$sing.event, function(x) pred_indiv_df$experiment[pred_indiv_df$sing.event == x][1])

coordination_results$group.size <- sapply(coordination_results$sing.event, function(x) length(unique(pred_indiv_df$indiv[pred_indiv_df$sing.event == x])))

coordination_results <- coordination_results[!is.na(coordination_results$p.value), ]

coordination_results <- coordination_results[!is.na(coordination_results$coor.score),]

# remove effect of group size
coordination_results$coor.score_cr <- residuals(lm(coor.score ~ group.size, coordination_results))

write.csv(coordination_results, "./data/processed/vocal_coordination_test_results.csv", row.names = FALSE)

```

```{r model measure coordination, eval = FALSE}

coordination_results <- read.csv("./data/processed/vocal_coordination_test_results.csv")



metap::sumz(1 - coordination_results$p.value,  weights = coordination_results$group.size)

metap::sumz(1 - coordination_results$p.value[coordination_results$experiment == "Regular"], weights = coordination_results$group.size[coordination_results$experiment == "Regular"])

metap::sumz(1 - coordination_results$p.value[coordination_results$experiment == "Artificial"], weights = coordination_results$group.size[coordination_results$experiment == "Artificial"])

mod <- brm(
          formula = coor.score ~ experiment + (1 | group.size),
          iter = iter,
          thin = 1,
          data = coordination_results,
          family = gaussian(),
          silent = 2,
          chains = chains,
          cores = chains,
          prior = c(prior(normal(0, 10), "b"), prior(student_t(3, 0, 20), "sigma")),
          file = "./data/processed/coordination_score_by_group_type",
          file_refit = "always"
              
          )

```

### Histogram of p values

```{r c coordination, eval = TRUE}
coordination_results <- read.csv("./data/processed/vocal_coordination_test_results.csv")

ggplot(coordination_results, aes(x = p.value)) + 
  geom_histogram(fill = viridis(10, alpha = 0.5)[3]) + 
  facet_grid(~ experiment)

```

#### Coordination scores
```{r c coordination 2, eval = TRUE}

agg_dat <- aggregate(coor.score ~ experiment, coordination_results, length)
agg_dat$n.labels <- paste("n groups =", agg_dat$coor.score)

# composed box plot
ggplot(coordination_results, aes(x = experiment, y = coor.score)) +
  ## add half-violin from {ggdist} package
  geom_hline(yintercept = 0, lty = 2) +
  ggdist::stat_halfeye(
    fill = fill_color,
    alpha = 0.5,
    ## custom bandwidth
    adjust = .5,
    ## adjust height
    width = .6,
    .width = 0,
    ## move geom to the cright
    justification = -.2,
    point_colour = NA
  ) +
  geom_boxplot(fill = fill_color,
    width = .15,
    ## remove outliers
    outlier.shape = NA ## `outlier.shape = NA` works as well
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    color = fill_color,
    ## draw jitter on the left
    side = "l",
    ## control range of jitter
    range_scale = .4,
    ## add some transparency
    alpha = .5,
    transformation = ggplot2::position_jitter(height = 0)
  ) +
    labs(x="Group", y= "Coordination score") +
  geom_text(data = agg_dat, aes(y = rep(-0.8, nrow(agg_dat)), x = c(1, 2), label = n.labels), nudge_x = 0, size = 6) 
  
```

### Combine p-values

Overall p-value:
```{r coordination 3, eval = TRUE}
metap::sumz(1 - coordination_results$p.value,  weights = coordination_results$group.size)

```

Coordination in real groups
```{r coordination 4, eval = TRUE}
metap::sumz(1 - coordination_results$p.value[coordination_results$experiment == "Regular"], weights = coordination_results$group.size[coordination_results$experiment == "Regular"])
```

Coordination in artificial groups
```{r coordination 5, eval = TRUE}

metap::sumz(1 - coordination_results$p.value[coordination_results$experiment == "Artificial"], weights = coordination_results$group.size[coordination_results$experiment == "Artificial"])

```

### Regression model comparing coordination score between treatments
```{r, results='asis'}

extended_summary(read.file = "./data/processed/coordination_score_by_group_type.rds", gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_", remove.intercepts = TRUE, highlight = TRUE, print.name = FALSE)

```

### Comparing treatments against zero
```{r, results='asis'}

contrasts(read.file = "./data/processed/coordination_score_by_group_type.rds", predictor = "experiment", gsub.pattern = "experiment", gsub.replacement = "", highlight = TRUE, plot = TRUE, non.zero = TRUE, html.table = TRUE)

```

&nbsp;

<div class="alert alert-info">

### Takeaways {.unnumbered .unlisted}

- No clear pattern of time partitioning in group flight calling

</div>

&nbsp;

## Acoustic space partitioning

- Use only groups in which the smallest individual sample size is higher than 4
- Measured as the pirwise density overlap between all the individuals within a group
- Rarefaction was used (with target n = 10) to control for unequal call sample sizes between individuals

```{r measure acoustic space, eval = FALSE}

ns <- lapply(acous_param_l, function(x) aggregate(selec ~ indiv + group + experiment, x, length))

n_indiv <- do_call(rbind, ns)

n_indiv <- n_indiv[n_indiv$experiment == "vuelo solo", ]

# keep groups with at least 5 calls per individual
n_indiv <- n_indiv[!n_indiv$group %in% unique(n_indiv$group[n_indiv$selec < 10]), ]

sub_acous_param_l <- acous_param_l[names(acous_param_l) %in% unique(n_indiv$group)]

acous_param <- do.call(rbind, sub_acous_param_l)

acous_param <- acous_param[acous_param$experiment == "vuelo solo", ]
acous_param$indiv.group <- paste(acous_param$indiv, acous_param$group, sep = "-")

# choose one set of calls per individual
acous_param <- acous_param[acous_param$indiv.group %in% sapply(unique(acous_param$indiv), function(x) acous_param$indiv.group[acous_param$indiv == x][1]), ]

umap_acous <- umap(acous_param[ , 3:222])

umap_acous$org.data <- acous_param

umap_dat <- data.frame(group = umap_acous$org.data$group, type = umap_acous$org.data$experiment, indiv = umap_acous$org.data$indiv, umap_acous$layout)

sp_sims_l <- pbapply::pblapply(names(sub_acous_param_l), function(x){
  
  indivs <- unique(sub_acous_param_l[[x]]$indiv[sub_acous_param_l[[x]]$experiment == "vuelo solo"])
  
  X <- umap_dat[umap_dat$indiv %in% indivs, ]

  sp_sim <- rarefact_space_similarity(X = X, dimensions = c("X1", "X2"), group = "indiv", parallel = 10, type = "mean.density.overlap", n = 10, iterations = 50, pb = FALSE)
  
  sp_sim <- sp_sim[, c("group.1", "group.2", "mean.overlap")]
  sp_sim$mean.distance <- rarefact_space_similarity(X = X, dimensions = c("X1", "X2"), group = "indiv", parallel = 10, type = "centroid.distance", n = 10, iterations = 50, pb = FALSE)$mean.distance
  
  sp_sim$group <- x
  sp_sim$type <- grep("gr",sub_acous_param_l[[x]]$experiment, value = TRUE)[1]
  names(sp_sim)[1:2] <- c("indiv.1", "indiv.2")
  sp_sim$group.size <- length(indivs)
  
  return(sp_sim)
})

sp_sims <- do.call(rbind, sp_sims_l)

sp_sims$experiment.type <- ifelse(sp_sims$type == "vuelo grupos mixtos", "Artificial", "Regular")

write.csv(sp_sims, "./data/processed/acoustic_similarity_between_individuals_by_group.csv", row.names = FALSE)

```

### Acoustic space overlap by group type and number of group size (facets)
```{r compare acoustic similarities plot, eval = TRUE, fig.height=5}

sp_sims <- read.csv("./data/processed/acoustic_similarity_between_individuals_by_group.csv")

sp_sims$experiment.type <- factor(ifelse(sp_sims$experiment.type == "Regular", "Regular", "Artificial"), levels = c("Regular", "Artificial"))

ggplot(sp_sims, aes(y = mean.overlap, x = experiment.type)) +
  ## add half-violin from {ggdist} package
  ggdist::stat_halfeye(
    fill = fill_color,
    alpha = 0.5,
    ## custom bandwidth
    adjust = .5,
    ## adjust height
    width = .6,
    .width = 0,
    ## move geom to the cright
    justification = -.2,
    point_colour = NA
  ) +
  geom_boxplot(fill = fill_color,
    width = .15,
    ## remove outliers
    outlier.shape = NA ## `outlier.shape = NA` works as well
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    color = fill_color,
    ## draw jitter on the left
    side = "l",
    ## control range of jitter
    range_scale = .4,
    ## add some transparency
    alpha = .5,
    transformation = ggplot2::position_jitter(height = 0)
  ) +
    labs(x="Group flight", y="Acoustic space overlap"
  ) +
   theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
   facet_wrap( ~ group.size)


```

```{r compare acoustic similarities model, eval = FALSE}
mod <- brm(
          formula = mean.overlap ~ experiment.type + experiment.type:group.size + (1 | group),
          iter = iter,
          thin = 1,
          data = sp_sims,
          family = gaussian(),
          silent = 2,
          chains = chains,
          cores = chains, 
          prior = priors, 
          file = "./data/processed/acoustic_similarity_between_individuals_by_group"
          )

```

### Regression model comparing overlap between group types
```{r  compare acoustic similarities results, eval = TRUE, results = "asis"}

extended_summary(read.file = "./data/processed/acoustic_similarity_between_individuals_by_group.rds", gsub.pattern = "b_experiment.type", gsub.replacement = "", remove.intercepts = TRUE, highlight = TRUE, print.name = FALSE)

```

&nbsp;

<div class="alert alert-info">

### Takeaways {.unnumbered .unlisted}

- No acoustic space partitioning in group flight calling (artificial or regular groups)

</div>

&nbsp;

```{r DEPRECATED Compare gaps duration among sexes, eval = FALSE, echo = FALSE}

str(gaps_indiv)

indivs <- unique(gaps_indiv$indiv) 
indivs <- indivs[indivs != "group"]

gaps_indiv_sex_l <- lapply(indivs, function(x){
  
  X <- gaps_indiv[gaps_indiv$indiv == x, ]
  X <- X[X$type == "indiv.group", ]
  
  if(any(!is.na(X$sex)))
    X$sex <- X$sex[!is.na(X$sex)][1]
  
    if(any(!is.na(X$age)))
    X$age <- X$age[!is.na(X$age)][1]
 
     if(any(!is.na(X$reprod.stg)))
    X$reprod.stg <- X$reprod.stg[!is.na(X$reprod.stg)][1]
  
  X <- X[!is.na(X$sex), ]
  
  return(X)
  }
)

gaps_indiv_sex <- do.call(rbind, gaps_indiv_sex_l)

table(gaps_indiv_sex$sex, gaps_indiv_sex$experiment)

ggplot(gaps_indiv_sex, aes(x = sex, y = gaps)) + 
  geom_violin() + 
  facet_wrap(~ experiment) +
  theme_classic()

```

```{r DEPRECATED Calls location in group flight by individual, eval = F, fig.height= 8, fig.width= 12, echo=FALSE}

# read as RDS
agg_pred <- readRDS("./data/processed/predicted_individual_in_group_flights_2020_&_2021.RDS")

agg_pred$orig.sound.files <- paste0(sapply(agg_pred$sound.files, function(x) strsplit(x, ".wav")[[1]][1]), ".wav")

agg_pred$start <- sapply(agg_pred$sound.files, function(x) seltab$orig.start[seltab$sound.files == x])

agg_pred$end <- sapply(agg_pred$sound.files, function(x) seltab$orig.end[seltab$sound.files == x])

agg_pred$middle.time <- (agg_pred$start + agg_pred$end) / 2
 
agg_pred$abbr_indiv <- as.numeric(as.factor(agg_pred$pred_indiv))

agg_pred$group2 <- make.names(agg_pred$group)


agg_pred <- agg_pred[!duplicated(agg_pred[, c("group", "abbr_indiv", "start", "end")]), ]

agg_pred$row <- 1:nrow(agg_pred)

# this are repeated
# agg_pred[c(200, 206, 156, 219, 30, 31, 9, 60, 160, 225, 486, 435, 438, 432), ]

agg_pred <- agg_pred[-c(200, 156, 30, 9, 160, 435, 432), ]

out <- lapply(unique(agg_pred$group2), function(x) {
  
  print(x)
    Y <- agg_pred[agg_pred$group2 == x, ]

    # jpeg(filename = file.path("img/group_call_through_time", paste0(x, ".jpeg")), width = 480 * 3, height = 480 * 2, res = 200)

    par(mar = c(1, 0, 0, 0), mfrow = c(4, 1))  
  
  Y$middle.time <- Y$middle.time - min(Y$middle.time) 
  
  num.ind <- as.numeric(as.factor(Y$abbr_indiv))
  
  num.ind <- scale(num.ind, center = TRUE) 
    
  # sq <- seq(0, max(Y$middle.time), length.out = 5)
  sq <- seq(0, 120, length.out = 5)
  
  
  for (i in 1:(length(sq) - 1))  {
  
    plot(0, 0, col = "white", xlim = c(sq[i] - 0.05, sq[i + 1] + 0.05), ylim = range(num.ind) + c(-2, 2), bty = "u")
    
  points(x = Y$middle.time, y = num.ind, cex = 7, col = viridis(max(as.numeric(as.factor(Y$abbr_indiv))), alpha = 0.5)[as.numeric(as.factor(Y$abbr_indiv))], pch = 20)
    
    # text(x = Y$middle.time, y = num.ind + rnorm(nrow(Y), sd = 0.8), Y$row, cex = 0.6, col = "black")
  }
    
# dev.off()
  
})


```

```{r DEPRECATED markovchain, eval = F, echo = FALSE}


lapply(unique(agg_pred$group), function(x) {
  
    Y <- agg_pred[agg_pred$group == x, ]

    mc <- markov.test(Y$abbr_indiv, type = "rank")
    
    return(mc)
}
)



seq <- simulateMarkovChain(5000, matrix(0.25, 4, 4), states=c("a","c","g","t"))

markov.test(seq)

rnd.seq <- sample(seq)
markov.test(rnd.seq)



sequenceMatr <- createSequenceMatrix(sequence, sanitize = FALSE)
mcFitMLE <- markovchainFit(data = sequence)

rnd.seq <- sample(sequence)

rnd <- createSequenceMatrix(rnd.seq, sanitize = FALSE)
mcFitMLE2 <- markovchainFit(data = rnd.seq)


mcFitBSP <- markovchainFit(data = sequence, method = "bootstrap", nboot = 5, name = "Bootstrap Mc")


sim.sq <- rep(c("a", "b", "c", "d"), 100)

verifyMarkovProperty(sim.sq)

sim.sq2 <- sample(c("a", "b", "c", "d"), size = 100, replace = TRUE)
verifyMarkovProperty(sim.sq2)


weatherStates <- c("sunny", "cloudy", "rain")

byRow <- TRUE

weatherMatrix <- matrix(data = c(0.70, 0.2, 0.1,0.3, 0.4, 0.3, 0.2, 0.45, 0.35), byrow = byRow, nrow = 3,dimnames = list(weatherStates, weatherStates))

mcWeather <- new("markovchain", states = weatherStates, byrow = byRow, transitionMatrix = weatherMatrix, name = "Weather")

mcIgraph <- as(mcWeather, "igraph")
plot(mcIgraph)

```


---

<font size="4">Session information</font>

```{r session info, echo=F, eval = TRUE}

sessionInfo()

```

