---
title: <center><font size="6"><b>Roost entry coordination</b></font></center>
subtitle: <center><font size="4"><b>Group flight coordination in Thyroptera</b></font></center>
author: <center><font size="4"><a href="https://marce10.github.io/">Marcelo Araya-Salas, PhD</a> & <a href="https://www.batcr.com/">Gloriana Chaverri, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---


```{r packages, message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE, include = FALSE}

# unload all non-based packages
out <- sapply(paste('package:', names(sessionInfo()$otherPkgs), sep = ""), function(x) try(detach(x, unload = FALSE, character.only = TRUE), silent = T))


## add 'developer/' to packages to be installed from github
x <- c("viridis", "ggplot2", "knitr", "readxl", "pbapply", "brms", "kableExtra", "ggridges", "tidybayes")

aa <- lapply(x, function(y) {
  
  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]
  
  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  devtools::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

```

```{r functions and options, eval = TRUE, echo = TRUE}

opts_knit$set(root.dir = "..")

opts_chunk$set(fig.width = 12, fig.height = 8)

theme_set(theme_classic(base_size = 34))

```


# Data description
```{r data description, eval = TRUE}

dat <- as.data.frame(read_excel(path = "./data/raw/Anexo 1_Entrada a refugios-2022.xlsx"))

dat$entry.time <- as.numeric(dat$`Tiempo real`)

dat <- dat[!is.na(dat$entry.time),]


# dat$entry.time
# 
# str(dat)

cat("Number of individuals per type:")
(table(dat$`Tipo grupo`))

cat("Number of individuals per type removing missing data (NAs and Inf:")
(table(dat$`Tipo grupo`[!is.infinite(dat$entry.time) & !is.na(dat$entry.time)]))


table(dat$Video[dat$`Tipo grupo` == "Individual"])

table(dat$Video[dat$`Tipo grupo` != "Individual"])

```

# Stats

## Entry coordination
### Entry time difference
```{r Entry time difference plot, eval = TRUE}
group_dat <- dat[!is.infinite(dat$entry.time) & !is.na(dat$entry.time) & dat$`Tipo grupo` != "Individual", ]

# get difference to first entry
group_dat_l <- lapply(unique(group_dat$Grupo), function(x){
  # print(x)
  X <- group_dat[group_dat$Grupo == x, ]
  X$entry.time.diff <- X$entry.time - min(X$entry.time)
  X <- X[-which.min(X$entry.time.diff), ]
  
  X$group.size <- if (nrow(X) > 0) nrow(X) - 1 else vector()
  return(X)
  })

group_dat <- do.call(rbind, group_dat_l)

group_dat$Group <- group_dat$Grupo
group_dat$entry.time.diff[group_dat$entry.time.diff == 0] <- 0.001
group_dat$type <- factor(ifelse(group_dat$`Tipo grupo` == "Mixto", "Aritficial", "Real"), levels = c("Aritficial", "Real"))

# individual plot solo vs individual in group
ggplot(group_dat, aes(x = type, y = log(entry.time.diff))) + 
  geom_boxplot() +
  # geom_point(size = 2) +
  # scale_color_viridis_c(alpha = 0.8) +
  labs(x = "Group type", y = "log(time after first entry (s))") + 
  # facet_grid( ~ experiment_f) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r Entry time difference stats, eval = FALSE}
chains <- 1
iter <- 5000

model_formulas <- c("entry.time.diff ~ type + (1 | Group) + (1 | group.size)", "entry.time.diff ~ 1 + (1 | Group) + (1 | group.size)")

brms_models <- lapply(model_formulas, function(x){

  mod <- brm(
          formula = x,
          iter = iter,
          thin = 1,
          data = group_dat,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains
          )
  
  mod <- add_criterion(mod, c("loo"))

  return(mod)
  })

names(brms_models) <- model_formulas

saveRDS(brms_models, "./data/processed/roost_entry_coordination_models.RDS")

```

```{r Entry time difference results, eval = TRUE, results = "asis"}

brms_models <- readRDS("./data/processed/roost_entry_coordination_models.RDS")

cat(paste(length(brms_models),"models evaluated:\n"))

for(i in names(brms_models))
  cat(paste("-  ", i, "\n"))

comp_mods <- loo_compare(brms_models[[1]], brms_models[[2]], model_names = names(brms_models))


df1 <- kbl(comp_mods, row.names = TRUE, escape = FALSE, format = "html", digits = 3)

cat("Compare models:")
  
 kable_styling(df1, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 12)

cat("Best model:\n")
cat(paste("-  ", rownames(comp_mods)[1], "\n"))

# best model
if (!grepl("1 +", rownames(comp_mods)[1], fixed = TRUE))
summary_brm_model(brms_models[[rownames(comp_mods)[1]]], gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_")

```

&nbsp;

<div class="alert alert-info">

#### Takeaways

-  No differences in roost entry coordination between real and artificial groups
</div>

&nbsp;

---

### Simulation entry time difference
```{r Simulation entry time difference, eval = FALSE}

group_dat <- dat[!is.infinite(dat$entry.time) & !is.na(dat$entry.time) & dat$`Tipo grupo` != "Individual", ]

# get difference to first entry
group_dat_l <- lapply(unique(group_dat$Grupo), function(x){
  # print(x)
  X <- group_dat[group_dat$Grupo == x, ]
  X$entry.time.diff <- X$entry.time - min(X$entry.time)
  X <- X[-which.min(X$entry.time.diff), ]
  
  X$group.size <- if (nrow(X) > 0) nrow(X) + 1 else vector()
  return(X)
  })

group_dat <- do.call(rbind, group_dat_l)



group_dat$Group <- group_dat$Grupo
group_dat$entry.time.diff[group_dat$entry.time.diff == 0] <- 0.001
group_dat$type <- factor(ifelse(group_dat$`Tipo grupo` == "Mixto", "Artificial", "Real"), levels = c("Simulated", "Artificial", "Real"))

# make random groups from individual flights
indiv_dat <- dat[!is.infinite(dat$entry.time) & !is.na(dat$entry.time) & dat$`Tipo grupo` == "Individual", ]

indivs <- unique(indiv_dat$Individuo)


sim_groups_l <- pblapply(1:1000, function(x){
  
 g_size <- sample(group_dat$group.size)
  
 n_group <- sum(cumsum(g_size) <= length(indivs))  
 
 g_size <- g_size[1:n_group]
 
 sampled_indivs <- sample(indivs, sum(g_size))
 
 indivs_split <- split(sampled_indivs, f = unlist(sapply(1:n_group, function(x) rep(x, g_size[x]))))
 
 sub_sim_groups_l <- lapply(1:length(indivs_split), function(y) {
   
   sim_group <- indiv_dat[indiv_dat$Individuo %in% indivs_split[[y]], ]
   sim_group$Group <- paste("sim",x, y, sep = "-")
 
   sim_group$entry.time.diff <- sim_group$entry.time - min(sim_group$entry.time)
   sim_group <- sim_group[-which.min(sim_group$entry.time.diff), ]
   sim_group$group.size <- nrow(sim_group) + 1
   sim_group$type <- factor("Simulated")
   
   out <- rbind(sim_group, group_dat)
   
   return(out)
   }
  )
 
 sub_sim_group <- do.call(rbind, sub_sim_groups_l)
 
 return(sub_sim_group)
})


sim_group_n <- sapply(sim_groups_l, function(x) length(unique(x$Group[x$type == "Simulated"])))

table(sim_group_n)

# keep those with same number of simulated groups
sim_groups_l <- sim_groups_l[which(sim_group_n == 5)]

sim_groups_l <- lapply(sim_groups_l, function(x){
  
  Y <- x[x$type == "Simulated", ]
  Y$Group <- as.numeric(as.factor(Y$Group)) 
  Y <- rbind(Y, x[x$type != "Simulated", ])
  
  return(Y)
  })

# keep only 30
sim_groups_l <- sim_groups_l[1:30]

sim_group_dat <- do.call(rbind, lapply(sim_groups_l, function(x) x[x$type == "Simulated", ]))

sim_group_dat <- rbind(sim_group_dat, group_dat)


# individual plot solo vs individual in group
ggplot(sim_group_dat, aes(x = type, y = log(entry.time.diff))) + 
  geom_boxplot() +
  # geom_point(size = 2) +
  # scale_color_viridis_c(alpha = 0.8) +
  labs(x = "Flight", y = "log(time after first entry (s))") + 
  # facet_grid( ~ experiment_f) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

saveRDS(sim_groups_l, "./data/processed/simulated_data_entry_time_difference.RDS")

```

```{r simulation stats, eval = FALSE}

sim_groups_l <- readRDS("./data/processed/simulated_data_entry_time_difference.RDS")

chains <- 1
iter <- 10000

mods <- brm_multiple(
          formula = entry.time.diff ~ type + (1 | Group) + (1 | group.size),
          iter = iter,
          thin = 1,
          data = sim_groups_l,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains,
          combine = FALSE
          )

combined_mod <- combine_models(mlist = mods, check_data = FALSE)


saveRDS(combined_mod, "./data/processed/simulated_data_entry_time_difference_models.RDS")

```

```{r simulations results, eval = TRUE, results = "asis"}

combined_mod <- readRDS("./data/processed/simulated_data_entry_time_difference_models.RDS")

summary_brm_model(combined_mod)

```

 Real vs artificial:
```{r simulation contrast}

contrasts <- c(real_vs_artificial = "typeReal - typeArtificial = 0")
hypothesis(combined_mod, contrasts)

```

&nbsp;

<div class="alert alert-info">

#### Takeaways

- Both real and artificial groups enter more sync'ed than simulated groups 
-  No differences in roost entry coordination between real and artificial groups
</div>

&nbsp;


## Entry efficiency
### Time to entry
```{r individual call rate plot, eval = TRUE}

group_dat <- dat[!is.infinite(dat$entry.time) & !is.na(dat$entry.time) & dat$`Tipo grupo` != "Individual", ]

aggregate(Video ~ Grupo, group_dat, function(x) length(unique(x)))


# get difference to first entry
group_dat_l <- lapply(unique(group_dat$Grupo), function(x){
  # print(x)
  X <- group_dat[group_dat$Grupo == x, ]
 
  if (length(unique(X$Video)) > 1) {
    
  tab <- table(X$Video)
  
  X <- X[X$Video == names(tab)[which.max(tab)], ]  
  }
  
  X$group.size <- if (nrow(X) > 0) nrow(X) else vector()
  
  return(X)
  })



group_dat <- do.call(rbind, group_dat_l)

aggregate(Video ~ Grupo, group_dat, function(x) length(unique(x)))


group_dat$Group <- group_dat$Grupo
group_dat$type <- factor(ifelse(group_dat$`Tipo grupo` == "Mixto", "Aritficial", "Real"), levels = c("Aritficial", "Real"))

# individual plot solo vs individual in group
ggplot(group_dat, aes(x = type, y = log(entry.time))) + 
  geom_boxplot() +
  # geom_point(size = 2) +
  # scale_color_viridis_c(alpha = 0.8) +
  labs(x = "Group type", y = "log(time to entry (s))") + 
  # facet_grid( ~ experiment_f) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r individual call rate stats, eval = FALSE}

chains <- 1
iter <- 5000

model_formulas <- c("entry.time ~ type + (1 | Group) + (1 | group.size)", "entry.time ~ 1 + (1 | Group) + (1 | group.size)")

brms_models <- lapply(model_formulas, function(x){

  mod <- brm(
          formula = x,
          iter = iter,
          thin = 1,
          data = group_dat,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains
          )
  
  mod <- add_criterion(mod, c("loo"))

  return(mod)
  })

names(brms_models) <- model_formulas

saveRDS(brms_models, "./data/processed/roost_entry_time_models.RDS")

```

```{r individual call rate results, eval = TRUE, results = "asis"}

brms_models <- readRDS("./data/processed/roost_entry_time_models.RDS")

cat(paste(length(brms_models),"models evaluated:\n"))

for(i in names(brms_models))
  cat(paste("-  ", i, "\n"))

comp_mods <- loo_compare(brms_models[[1]], brms_models[[2]], model_names = names(brms_models))


df1 <- kbl(comp_mods, row.names = TRUE, escape = FALSE, format = "html", digits = 3)

cat("Compare models:")
  
 kable_styling(df1, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 12)

cat("Best model:\n")
cat(paste("-  ", rownames(comp_mods)[1], "\n"))

# best model
if (!grepl("1 +", rownames(comp_mods)[1], fixed = TRUE))
summary_brm_model(brms_models[[rownames(comp_mods)[1]]], gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_")

```

### Simulation entry time difference
```{r Simulation entry time, eval = FALSE}

group_dat$type <- factor(ifelse(group_dat$`Tipo grupo` == "Mixto", "Artificial", "Real"), levels = c("Simulated", "Artificial", "Real"))

# make random groups from individual flights
indiv_dat <- dat[!is.infinite(dat$entry.time) & !is.na(dat$entry.time) & dat$`Tipo grupo` == "Individual", ]

indivs <- unique(indiv_dat$Individuo)


sim_groups_l <- pblapply(1:1000, function(x){
  
 g_size <- sample(group_dat$group.size)
  
 n_group <- sum(cumsum(g_size) <= length(indivs))  
 
 g_size <- g_size[1:n_group]
 
 sampled_indivs <- sample(indivs, sum(g_size))
 
 indivs_split <- split(sampled_indivs, f = unlist(sapply(1:n_group, function(x) rep(x, g_size[x]))))
 
 sub_sim_groups_l <- lapply(1:length(indivs_split), function(y) {
   
   sim_group <- indiv_dat[indiv_dat$Individuo %in% indivs_split[[y]], ]
   sim_group$Group <- paste("sim", x, y, sep = "-")
   sim_group$group.size <- nrow(sim_group)
   sim_group$type <- factor("Simulated")
   
   out <- rbind(sim_group, group_dat)
   
   return(out)
   }
  )
 
 sub_sim_group <- do.call(rbind, sub_sim_groups_l)
 
 return(sub_sim_group)
})


sim_group_n <- sapply(sim_groups_l, function(x) length(unique(x$Group[x$type == "Simulated"])))

table(sim_group_n)

# keep those with same number of simulated groups
sim_groups_l <- sim_groups_l[which(sim_group_n == 5)]

sim_groups_l <- lapply(sim_groups_l, function(x){
  
  Y <- x[x$type == "Simulated", ]
  Y$Group <- as.numeric(as.factor(Y$Group)) 
  Y <- rbind(Y, x[x$type != "Simulated", ])
  
  return(Y)
  })

# keep only 30
sim_groups_l <- sim_groups_l[1:30]

sim_group_dat <- do.call(rbind, lapply(sim_groups_l, function(x) x[x$type == "Simulated", ]))

sim_group_dat <- rbind(sim_group_dat, group_dat)


# individual plot solo vs individual in group
ggplot(sim_group_dat, aes(x = type, y = log(entry.time))) + 
  geom_boxplot() +
  # geom_point(size = 2) +
  # scale_color_viridis_c(alpha = 0.8) +
  labs(x = "Flight", y = "log(entry time (s))") + 
  # facet_grid( ~ experiment_f) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

saveRDS(sim_groups_l, "./data/processed/simulated_data_entry_time.RDS")

```


```{r simulation time stats, eval = FALSE}

sim_groups_l <- readRDS("./data/processed/simulated_data_entry_time.RDS")

chains <- 1
iter <- 10000

mods <- brm_multiple(
          formula = entry.time ~ type + (1 | Group) + (1 | group.size),
          iter = iter,
          thin = 1,
          data = sim_groups_l,
          family = lognormal(),
          silent = 2,
          chains = chains,
          cores = chains,
          combine = FALSE
          )

combined_mod <- combine_models(mlist = mods, check_data = FALSE)


saveRDS(combined_mod, "./data/processed/simulated_data_entry_time_models.RDS")

```

```{r simulations time results, eval = TRUE, results = "asis"}

combined_mod_time <- readRDS("./data/processed/simulated_data_entry_time_models.RDS")

summary_brm_model(combined_mod_time)

```

 Real vs artificial:
```{r simulation time contrast}

contrasts <- c(real_vs_artificial = "typeReal - typeArtificial = 0")
hypothesis(combined_mod_timeda, contrasts)

```

---

## Propensity to enter the roost

```{r propensity to enter leaves plot, eval = TRUE}
group_dat <- dat[dat$`Tipo grupo` != "Individual", ]
group_dat$indiv.entry <- ifelse(is.na(group_dat$entry.time) | is.infinite(group_dat$entry.time), 0, 1)


group_dat$Group <- group_dat$Grupo
group_dat$type <- factor(ifelse(group_dat$`Tipo grupo` == "Mixto", "Aritficial", "Real"), levels = c("Aritficial", "Real"))  

agg_dat <- aggregate(indiv.entry ~ Group + type, group_dat, FUN = sum)
agg_dat$group.size <- aggregate(indiv.entry ~ Group + type, group_dat, FUN = length)$indiv.entry


agg_dat$entry <- "Partial"
agg_dat$entry[agg_dat$group.size == agg_dat$indiv.entry] <- "Full"
agg_dat$entry[agg_dat$indiv.entry == 0] <- "None"

count_dat <- aggregate(Group ~ entry + type, agg_dat, length)
count_dat$entry <- factor(count_dat$entry, levels = c("None", "Partial", "Full"))

# Change barplot fill colors by groups
ggplot(count_dat, aes(x=entry, y=Group, fill=type)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  scale_fill_viridis_d(begin = 0.3, end = 0.8) +
  theme_classic()


```

```{r propensity to enter leaves stats, eval = FALSE}
# mean centering
agg_dat$group.size <- agg_dat$group.size - mean(agg_dat$group.size)

chains <- 1
iter <- 10000

agg_dat$entry[agg_dat$entry == "None"] <-  "1_None"

model_formulas <- c("entry ~ type + (1 | Group) + ( 1 | group.size)")

brms_models <- lapply(model_formulas, function(x){
print(x)
  mod <- brm(data = agg_dat, 
            family = categorical, 
            formula = x,
            iter = iter,
            chains = chains, 
            cores = chains
              )
  
  try(mod <- add_criterion(mod, c("loo")))

  return(mod)
  })

names(brms_models) <- model_formulas

saveRDS(brms_models, "./data/processed/roost_entry_models.RDS")

```

```{r, results = "asis"}
brms_models <- readRDS("./data/processed/roost_entry_models.RDS")

summary_brm_model(brms_models$`entry ~ type + (1 | Group) + ( 1 | group.size)`)
```


```{r}
contrasts <- c(one = "muFull_Intercept - muFull_typeReal = 0")
hypothesis(brms_models$`entry ~ type + (1 | Group) + ( 1 | group.size)`, contrasts)

```

```{r propensity to enter leaves results, eval = FALSE, results = "asis"}

brms_models <- readRDS("./data/processed/roost_entry_models.RDS")

cat(paste(length(brms_models),"models evaluated:\n"))

for(i in names(brms_models))
  cat(paste("-  ", i, "\n"))

comp_mods <- loo_compare(brms_models[[1]], brms_models[[2]],  brms_models[[3]], model_names = names(brms_models))


df1 <- kbl(comp_mods, row.names = TRUE, escape = FALSE, format = "html", digits = 3)

cat("Compare models:")
  
 kable_styling(df1, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 12)

cat("Best model:\n")
cat(paste("-  ", rownames(comp_mods)[1], "\n"))

# best model
if (!grepl("1 +", rownames(comp_mods)[1], fixed = TRUE))
  summary_brm_model(brms_models[[rownames(comp_mods)[1]]], gsub.pattern = "experiment.type", gsub.replacement = "solo_vs_")

```

Full entry of real vs artificial groups:
```{r propensity results contrasts, eval = TRUE}


```

```{r, eval = FALSE}

real_distances_l <- readRDS("./data/processed/real_trajectory_mean_distances.RDS")

```

&nbsp;

<div class="alert alert-info">

### Takeaways

-  No clear differences between real and artificial groups in propensity to enter roost (full group entry)
</div>

&nbsp;




---


<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```
